---
title: 'Case Study 00: This is a template'
author: "Author 1 and Author 2"
date: "March 00, 2015"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
---

```{r setup,results='hide',warning=FALSE,echo=FALSE}
# A few initial definitions just to make sure all required packages are installed. Change as needed.
# NOTE: It may echo some weird messages to the PDF on the first compile (package installation messages). Run twice and the problem will (hopefully) go away.
if (!require(ggplot2, quietly = TRUE)){
      install.packages("ggplot2")
      }
if (!require(devtools, quietly = TRUE)){
      install.packages("devtools")
      }
 if (!require(broom, quietly = TRUE)){
       devtools::install_github("dgrtwo/broom")
      }
if (!require(GGally, quietly = TRUE)){
      install.packages("GGally")
      }
```
## Descrição do Problema
O IMC é uma medida utilizada na área de saúde como um indicador simples capaz de encontrar correlações entre o peso e altura do paciente com doenças decorrentes de obesidade, sendo inclusive capaz de classificar em diferentes grupos para obter maior precisão na análise em questão [ADD. REF].

Apesar dos problemas decorrentes da simplicidade do teste [ADD REF.] e métodos mais robustos tais como a bioimpedância [ADD REF.] estarem presentes, continua sendo uma medida interessante pela facilidade e baixo custo de encontrá-la. A partir de tal importância médica, muitas questões naturalmente surgem relacionadas a testes entre diferentes populações.

Neste estudo, serão propostos três testes de forma a compreender possíveis relações e diferenças presentes na amostragem realizada nos anos de 2016 e 2017 no Departamento de Engenharia da UFMG. Os testes propostos são:

* Diferença entre IMC de Homens e Mulheres;
* Diferença de IMC entre os dois anos em estudo;
* Diferença entre alunos de Graduação e Pós-Graduação em 2017.

Por meio destes testes, podemos obter informações relevantes capazes de compreender diferenças baseadas em idade, sexo e ano de estudo.

## Design dos Experimentos

### 1. The easy way
Use [RStudio](http://rstudio.com/) as your editor, open the **.Rmd** file and click the _Knit PDF_ button at the top of the editor.

### 2. The slightly-less-easy way
If you're using any other R editor (such as the basic [R](http://cran.r-project.org) editor), you have to use the _render()_ function from the **rmarkdown** package:

```{r, eval=F}
install.packages("devtools")                    # you only have to install it once
library(devtools)
install_github("rstudio/rmarkdown")             # you only have to install it once
library(rmarkdown)
render("report_template.Rmd","pdf_document")    # this renders the pdf
```

## Design dos Experimentos

### Diferença entre IMC de Homens e Mulheres
Dada a suposta natureza 
$$\begin{cases} H_0: \mu = 10&\\H_1: \mu<10\end{cases}$$

including the reasons behind your choices of the value for $H_0$ and the directionality (or not) of $H_1$. 

This is also the place where you should discuss (whenever necessary) your definitions of minimally relevant effects ($\delta^*$), sample size calculations, choice of power and significance levels, and any other relevant information about specificities in your data collection procedures.

### Description of the data collection
Whenever needed, you can also include an (optional) subsection describing the actual data collection, how it was performed, any adaptations or unexpected events that may have occurred, etc. Subsections like this can also be used for the sample size calculations or any other aspect that requires a longer discussion.

## Experimentos
O primeiro passo é carregar os dados e calcular o IMC das amostras.

```{r loaddata}
imc2016 <- read.csv('imc_20162.csv')
imc2017 <- read.csv('CS01_20172.csv', sep = ';')

# Filtragem de dados
imc2016 <- imc2016[imc2016$Course == 'PPGEE',]

# Padronização do dataframe
colnames(imc2016)[colnames(imc2016)=='Gender'] <- 'Sex'
colnames(imc2016)[colnames(imc2016)=='Height.m'] <- 'height.m'
colnames(imc2016)[colnames(imc2016)=='Weight.kg'] <- 'weight.kg'
colnames(imc2017)[colnames(imc2017)=='Weight.kg'] <- 'weight.kg'
imc2016 = imc2016[,c('Sex', 'height.m', 'weight.kg')]
imc2017 = imc2017[,c('Sex', 'height.m', 'weight.kg')]
imc2016$amostra <- '2016'
imc2017$amostra <- '2017'

# Calculando o IMC de cada população
imc2016$imc <- imc2016$weight.kg/(imc2016$height.m^2)
imc2017$imc <- imc2017$weight.kg/(imc2017$height.m^2)

# União dos dataframes
imc <- rbind(imc2016, imc2017)

# Segregação de dados entre Homens e Mulheres
imc_feminino <- imc[imc$Sex == 'F',]
imc_masculino <- imc[imc$Sex == 'M',]

```

Em seguida, executamos o teste de Fligner-Killeen para verificar se a premissa 
de que as variâncias são iguais se aplica no experimento.

```{r fligner test}
fligner.test(imc ~ Sex, data = imc) 
```

Em seguida, executamos o teste de Shapiro-Wilker para verificar a premissa de 
normalidade das amostras

```{r shapiro test}
shapiro.test(imc$imc[imc$Sex == "F"]) 
shapiro.test(imc$imc[imc$Sex == "M"]) 

```

Com as premissas validadas, executamos o t-test:

```{r t-test homens e mulheres}
t.test(imc$imc ~ imc$Sex, 
       alternative = "less", 
       mu          = 0, 
       var.equal   = TRUE, 
       conf.level  = 0.95)
```

Finalmente, identificamos o poder do teste

```{r poder do teste}
power.t.test(delta       = 2.5, # mínimo que faz a mudança de faixa de IMC
             sd          = sd(imc$imc),
             sig.level   = 0.05,
             n = length(imc$imc),
             type        = "two.sample",
             alternative = "one.sided")
```


## Statistical Analysis
Your statistical analysis should come here. This is the place where you should fit your statistical model, get the results of your significance test, your effect size estimates and confidence intervals.

```{r fitmodel,results='hold'}
model<-aov(mpg~am*disp,data=mtcars)
summary(model)
```

### Checking Model Assumptions
The assumptions of your test should also be validated, and possible effects of violations should also be explored.
```{r resplots,fig.width=8,echo=TRUE,message=FALSE, fig.cap = "Residual plots for the anova model"}
par(mfrow=c(2,2), mai=.3*c(1,1,1,1))
plot(model,pch=16,lty=1,lwd=2)
```

### Conclusions and Recommendations
The discussion of your results, and the scientific/technical meaning of the effects detected, should be placed here. Always be sure to tie your results back to the original question of interest!